Early Observations of the Type Ia Supernova \abc:\\ \textbf{a Case of Interaction with Nearby, Unbound Material and/or} Strong Ejecta Mixing

### 2017-11-16 doc/ms.tex
update text in observations section

During the spring of 2016, the iPTF survey observed the field of \abc\ every night during dark time in either the $g_\mathrm{PTF}$ or 
$R_\mathrm{PTF}$ band.\footnote{P48 observations of \abc\ are reported in the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters throughout, which are similar to the SDSS $g'$ and Mould-$R$ filters, respectively (see \citealt{2012PASP..124..854O} for details on PTF calibration). The correction from the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ filters to SDSS $g'$ and $r'$ requires knowledge of the intrinsic source color (see Eqns.~1 and 2 in \citealt{2012PASP..124..854O}). The spectral diversity of SNe Ia in the days after explosion is poorly constrained, and as a result the color terms for \abc\ at these epochs are unknown. We proceed by assuming the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$ calibration is on the AB system, which strictly speaking is incorrect, but this does not fundamentally alter any of our conclusions.} 
Survey observations were conducted with the
CFH12K camera \citep{2008SPIE.7014E..4YR} on the Palomar Observatory 
48-inch telescope (P48; \citealt{2009PASP..121.1395L}). Images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN 
\citep{2017PASP..129a4002M}. The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.48\textwidth]{logLC_with_colors.pdf}
  \caption{The $g$-band
  light curve of \abc. Observations from different 
  telescopes are shown with different symbols. Time is measured in 
  rest-frame days relative to the time of first light, $t_0$ (see 
  \S\ref{sec:lc_fit}). Note that the horizontal axis is shown with 
  a linear scale for $-2 \; \mathrm{d} \le t - t_0 \le 3 \; 
  \mathrm{d}$ and a $\log$ scale for $t - t_0 > 3 \; \mathrm{d}$. 
  The black ticks near the
  bottom of the panel show epochs of spectroscopic observations.
  }
  \label{fig:lightcurve}
\end{figure}

After discovery, photometric observations in the $g'$, $r'$ and $i'$
filters were obtained with the SED Machine 
(SEDm; Blagorodnova et al.\ 2017, in prep.) mounted on the Palomar Observatory 
60-inch telescope (P60). We utilized the Fremling Automated Pipeline (\texttt{FPipe}; \citealt{2016A&A...593A..68F}) to subtract galaxy light from the SEDm images using archival Sloan Digital Sky Survey (SDSS) images as a reference. This pipeline then performed forced-PSF photometry at the location of \abc, which is calibrated to the SDSS catalog \citep{2014ApJS..211...17A}.

Photometric observations in the $BVg'r'i'$ filters were conducted by the Las Cumbres Observatory (LCO) 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{g'r'i'}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

The Reionization and Transients InfraRed (RATIR) camera on the autonomous 1.5\,m Harold L. Johnson Telescope at the Observatorio Astron\'{o}mico Nacional \citep{2012SPIE.8446E..10B,2012SPIE.8444E..5LW} was used to observe \abc\ in the $r'i'ZYJH$ filters. By design, RATIR lacks a cold shutter, which means IR dark frames are not available. Laboratory testing, however, confirms that the dark current is negligible in both IR detectors \citep{2012SPIE.8453E..1OF}.

@@

Spectroscopic observations of \abc\ were taken with a variety
of telescopes and instruments over multiple epochs spanning from a
couple of days after discovery to two months after $B$-band maximum. 
An observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}/\texttt{Matlab}. The optical
spectral evolution of \abc\ is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

@@

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.69\textwidth]{spectra.pdf}
  \caption{Observed spectral sequence of \abc. The spectra are 
  normalized by their median flux between 6,000 and 
  7,000$\,\textrm{\AA}$.  The phase of each spectrum 
  relative to \textbf{the time of $B$-band maximum} is shown. Telluric absorption
  bands are grayed out. Line identifications are provided for 
  the spectral features discussed in the text. 
  For clarity, high-resolution spectra obtained with the VLT have been omitted (see Ferretti et al.\ 2017, submitted, for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

@@

A detailed study of the reddening towards \abc\ is presented in a 
companion paper (Ferretti et al.\ 2017, submitted). Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in Ferretti et al.\ (2017) show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying significant extinction (e.g., \citealt{2012MNRAS.426.1465P}), Ferretti et al.\ compare the evolution of \abc\ to the well-observed normal Type Ia SN\,2011fe and only find evidence for a small amount of extinction. The empirical relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and \citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a poor tracer of reddening in Type Ia SNe. Thus, we adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ (Ferretti et al.\ 2017). For the remainder of our analysis we assume a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$.

>>>>

During the spring of 2016, the iPTF survey observed the field of \abc\ every
night during dark time in either the $g_\mathrm{PTF}$ or $R_\mathrm{PTF}$
band.\footnote{P48 observations of \abc\ are reported in the $g_\mathrm{PTF}$
and $R_\mathrm{PTF}$ filters throughout, which are similar to the SDSS $g'$
and Mould-$R$ filters, respectively (see \citealt{2012PASP..124..854O} for
details on PTF calibration). The correction from the $g_\mathrm{PTF}$ and
$R_\mathrm{PTF}$ filters to SDSS $g'$ and $r'$ requires knowledge of the
intrinsic source color (see Eqns.~1 and 2 in \citealt{2012PASP..124..854O}).
The spectral diversity of SNe Ia in the days after explosion is poorly
constrained, and as a result the color terms for \abc\ at these epochs are
unknown. We proceed by assuming the $g_\mathrm{PTF}$ and $R_\mathrm{PTF}$
calibration is on the AB system, which strictly speaking is incorrect, but
this does not fundamentally alter any of our conclusions.} Survey observations
were conducted with the CFH12K camera \citep{2008SPIE.7014E..4YR} on the
Palomar Observatory 48-inch telescope (P48; \citealt{2009PASP..121.1395L}).
Images were processed by the IPAC image-subtraction pipeline, which subtracts
background galaxy light \textbf{using deep} pre-SN images, and performs forced
point-spread-function (PSF) photometry at the location of the SN
\citep{2017PASP..129a4002M}. The photometry is then calibrated to the PTF
photometric catalog \citep{2012PASP..124..854O}.

\begin{figure}[htb] 
    \centering
    \includegraphics[width=0.48\textwidth]{logLC_with_colors.pdf} 
    \caption{
    The $g$-band light curve of \abc. Observations from different telescopes
    are shown with different symbols. Time is measured in rest-frame days
    relative to $t_0$ (see \S\ref{sec:lc_fit}). Note that the horizontal axis
    is shown with a linear scale for $-2 \; \mathrm{d} \le t - t_0 \le 3 \;
    \mathrm{d}$ and a $\log$ scale for $t - t_0 > 3 \; \mathrm{d}$.
    \textbf{Vertical} black ticks show epochs of spectroscopic observations. }
    \label{fig:lightcurve} 
\end{figure}

After discovery, $g'$, $r'$ and $i'$-band \textbf{photometry was} obtained
with the SED Machine (SEDm; \citealt{2017arXiv171002917B}) mounted on the
Palomar Observatory 60-inch telescope (P60). We utilized the Fremling
Automated Pipeline (\texttt{FPipe}; \citealt{2016A&A...593A..68F}) to subtract
galaxy light from the SEDm images using archival Sloan Digital Sky Survey
(SDSS) images as a reference. This pipeline then performed forced-PSF
photometry at the location of \abc, which is calibrated to the SDSS catalog
\citep{2014ApJS..211...17A}.

\textbf{The Las Cumbres Observatory (LCO) 1-m telescope network obtained
$BVg'r'i'$ photometry.} PSF photometry was measured on these images using the
\texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The \textit{BV}
magnitudes are calibrated to the Fourth USNO CCD Astrograph Catalog
\citep{2013AJ....145...44Z}, and the \textit{g'r'i'} magnitudes are calibrated
to SDSS Data Release 6 \citep{2008ApJS..175..297A}.

The Reionization and Transients InfraRed (RATIR) camera on the autonomous
1.5\,m Harold L. Johnson Telescope
\citep{2012SPIE.8446E..10B,2012SPIE.8444E..5LW} was used to observe \abc\ in
the $r'i'ZYJH$ filters. By design, RATIR lacks a cold shutter, which means IR
dark frames are not available. Laboratory testing, however, confirms that the
dark current is negligible in both IR detectors \citep{2012SPIE.8453E..1OF}.

@@

Spectroscopic observations of \abc\ were taken with a variety of telescopes
and instruments over multiple epochs \textbf{beginning $\sim$2\,d} after
discovery \textbf{and ending $\sim$2} months after $B$-band maximum. An
observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were
reduced using standard procedures in
\texttt{IDL}/\texttt{Python}/\texttt{Matlab}. The optical spectral evolution
of \abc\ is illustrated in Figure \ref{fig:spec_seq}, which excludes
high-resolution Very Large Telescope (VLT) spectra for clarity.

@@

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.69\textwidth]{spectra.pdf}
  \caption{
  Observed spectral sequence of \abc. The spectra are normalized by their
  median flux between 6000 and 7000$\,\textrm{\AA}$. The phase of each
  spectrum relative to \textbf{the time of $B$-band maximum} is shown.
  Telluric absorption bands are grayed out. Line identifications are provided
  for the spectral features discussed in the text. For clarity,
  high-resolution spectra obtained with the VLT have been omitted (see
  \citealt{2017A&A...606A.111F} for a detailed discussion of these spectra).}
  \label{fig:spec_seq}
\end{figure*}

@@

A detailed study of the reddening towards \abc\ is presented in a 
companion paper \citep{2017A&A...606A.111F}. Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in \citet{2017A&A...606A.111F} show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying significant extinction (e.g., \citealt{2012MNRAS.426.1465P}), \citeauthor{2017A&A...606A.111F} compare the evolution of \abc\ to the well-observed normal Type Ia SN\,2011fe and only find evidence for a small amount of extinction. The empirical relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and \citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a poor tracer of reddening in Type Ia SNe. Thus, we adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ \citep{2017A&A...606A.111F}. For the remainder of our analysis we assume a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$.
